---
title: "Introduction to tqr package"
author: "Mitsuo Shiota"
date: "2019-07-19"
output: 
  github_document:
    toc: TRUE
editor_options: 
  
  chunk_output_type: console
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/mitsuoxv/tqr/workflows/R-CMD-check/badge.svg)](https://github.com/mitsuoxv/tqr/actions)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

Updated: `r Sys.Date()`

# tqr: add-on to tsibble, inspired by tidyquant

When I utilize [`tidyquant` package](https://cran.r-project.org/web/packages/tidyquant/readme/README.html) for economic analyses, I often find myself writing similar functions repeatedly. I also find the idea of time aware dataframe of [`tsibble` package](https://cran.r-project.org/web/packages/tsibble/index.html) may help me deal with time series data more confidently. So I have decided to build this package to facilitate my work.

`tqr` package adds functions:

1. tq_diff: calculate differences
1. tq_ma: calculate moving averages
1. tq_gr: calculate growth rates
1. tq_sa: calculate seasonally adjusted values
1. cal_factory: function factory for calculation
1. cal_factory_zoo: function factory for calculation utilizing zoo package
1. cal_factory_ts: function factory for calculation utilizing ts class
1. cal_factory_xts: function factory for calculation utilizing xts package


## Installation

This package is only for my own use for now. "You" means "future me".

You can install the development version with:

``` r
remotes::install_github("mitsuoxv/tqr")

```

## Libraries

```{r setup, message=FALSE}
library(tidyverse)
library(tsibble)
library(tqr)

```
## Create time aware dataframe of tsibble class

I use monthly effective exchange rate indices of 60 countries, which BIS publishes on [its site](https://www.bis.org/statistics/eer.htm). `symbol` column is "reer" (real effective exchange rate), "neer" (nominal effective exchange rate) or "deflator", which I calculated as "neer" / "reer" * 100. I prepared `eer` as a tibble.

```{r read_eer}
eer

class(eer)

```

I transform `eer` from a tibble to a tsibble (`tbl_ts` class) by making clear that `date` is monthly, and by specifying `key` (category of values) and `index` (time pointing column). Once transformed, I can see meta info of a tsibble, like interval.

```{r as_tsibble}
eer_ts <- eer %>% 
  mutate(date = yearmonth(date)) %>% 
  as_tsibble(key = symbol, index = date)

class(eer_ts)

interval(eer_ts)

is_regular(eer_ts)

is_ordered(eer_ts)

```

## Insert missing rows, if necessary

Sometimes data miss some rows. For example, `corrupt_data` lacks "1994 3" and some other rows. This is dangerous, as my functions often use `lag`.

You can check if there are missing rows, and fill missing rows. Please refer to [`tsibble` package vignette "Handle implicit missingness with tsibble"](https://cran.r-project.org/web/packages/tsibble/vignettes/implicit-na.html).

In this case, `eer_ts` does not miss any rows. I can safely apply my functions.

```{r has_gaps}
has_gaps(eer_ts)

corrupt_data <- eer_ts[c(-3, -256, -900), ]

has_gaps(corrupt_data)

filled_data <- fill_gaps(corrupt_data)

filled_data

has_gaps(filled_data)


```

## tq_diff: calculate differences

Let us see the differences year-over-year.

```{r tq_diff, fig.width=8, fig.height=4}
eer_ts %>% 
  tq_diff(n = 12) %>% 
  pivot_longer(!c(date, symbol), names_to = "area") %>% 
  filter(symbol == "reer") %>% 
  filter(area %in% c("Japan", "Euro area", "United States")) %>% 
  ggplot(aes(x = date, y = value, color = area)) +
  geom_hline(yintercept = 0, size = 2, color = "white") +
  geom_line() +
  labs(
    title = "REER, differences year-over-year",
    x = NULL, y = NULL, color = NULL
  )

```

## tq_ma: calculate moving averages

Let us see 6 month moving average movements.

```{r tq_ma, fig.width=8, fig.height=4}
eer_ts %>% 
  tq_ma(n = 6) %>% 
  pivot_longer(!c(date, symbol), names_to = "area") %>% 
  filter(symbol == "reer") %>% 
  filter(area %in% c("Japan", "Euro area", "United States")) %>% 
  ggplot(aes(x = date, y = value, color = area)) +
  geom_hline(yintercept = 100, size = 2, color = "white") +
  geom_line() +
  labs(
    title = "REER, 6 month moving averages",
    x = NULL, y = NULL, color = NULL
  )

```

## tq_gr: calculate growth rates

Let us see year-over-year growth rates, percents.

```{r tq_gr, fig.width=8, fig.height=4}
eer_ts %>% 
  tq_gr(n = 12) %>% 
  pivot_longer(!c(date, symbol), names_to = "area") %>% 
  filter(symbol == "reer") %>% 
  filter(area %in% c("Japan", "Euro area", "United States")) %>% 
  ggplot(aes(x = date, y = value, color = area)) +
  geom_hline(yintercept = 0, size = 2, color = "white") +
  geom_line() +
  labs(
    title = "REER, year-over-year growth rates",
    x = NULL, y = NULL, color = NULL
  )

```

## Convert to lower frequency

You can convert from "month" to "quarter".

```{r convert_freq, fig.width=8, fig.height=4}
eer_q <- eer_ts %>% 
  pivot_longer(!c(date, symbol), names_to = "area") %>% 
  group_by(symbol, area) %>% 
  index_by(quarter = yearquarter(date)) %>% 
  summarize(value = mean(value))

eer_q %>% 
  filter(symbol == "reer") %>% 
  filter(area %in% c("Japan", "Euro area", "United States")) %>% 
  ggplot(aes(x = quarter, y = value, color = area)) +
  geom_hline(yintercept = 100, size = 2, color = "white") +
  geom_line() +
  labs(
    title = "REER, quarters",
    x = NULL, y = NULL, color = NULL
  )

```

## tq_sa: calculate seasonally adjusted values

Greece has seasonality in deflator which is nominal divided by real effective exchange rate.

Let us get seasonally adjust values. `tq_sa` utilizes [`seasonal` package](https://www.rdocumentation.org/packages/seasonal/versions/1.7.0).

```{r tq_sa, fig.width=8, fig.height=4}
greece <- eer_ts %>% 
  pivot_longer(!c(date, symbol), names_to = "area") %>% 
  filter(symbol == "deflator", area == "Greece") %>% 
  mutate(symbol = "original")

greece_sa <- greece %>% 
  tq_sa() %>% 
  mutate(symbol = "seasonally adjusted")

greece %>% 
  bind_rows(greece_sa) %>% 
  ggplot(aes(x = date, y = value, color = symbol)) +
  geom_hline(yintercept = 100, size = 2, color = "white") +
  geom_line() +
  labs(
    title = "Deflator, Greece",
    x = NULL, y = NULL, color = NULL
  )

```

## cal_factory: function factory for calculation

`tq_diff`, `tq_ma` and `tq_gr` functions are manufactured by `cal_factory` function. You can manufacture your own function. For example, here I manufacture a function to pick up every 6 month values.

```{r cal_factory, fig.width=8, fig.height=4}
tq_by6 <- cal_factory(
  function(num) {
    num[seq(1, length(num), by = 6)]
  },
  function(idx) {
    idx[seq(1, length(idx), by = 6)]
  },
  function(itv) {
    interval <- itv2list(itv)
    
    interval$number <- interval$number * 6

    paste0(interval$number, interval$category)
  }
)

greece %>% 
  tq_by6() %>% 
  ggplot(aes(x = date, y = value, color = symbol)) +
  geom_hline(yintercept = 100, size = 2, color = "white") +
  geom_line() +
  labs(
    title = "Deflator, every 6 month, Greece",
    x = NULL, y = NULL, color = NULL
  )

```

If you don't mind dirts and are willing to clean up later, you can manufacture a function that makes "date" column non dates. For example, I manufacture a function to calculate range.

```{r dirty_example1}
tq_range <- cal_factory(
  function(num) {
    range(num)
  },
  function(idx) {
    c("min", "max")
  },
  function(itv) {
    "?"
  }
)

eer_ts %>% 
  tq_range() %>% 
  pivot_longer(!c(date, symbol), names_to = "area") %>% 
  pivot_wider(names_from = date) %>% 
  mutate(range = max - min) %>% 
  filter(symbol == "deflator") %>% 
  arrange(desc(range))

```

Another dirty example. I calculate deflators' rates of change from the prior month, look at auto correlation of Algeria, and find high correlation every 12 months, i.e. seasonality. Then, I manufacture `tq_acf` function to get auto correlation values, and find that Greece has the highest seasonality.

```{r dirty_example2, fig.width=8, fig.height=4}
defl_gr1 <- eer_ts %>% 
  filter(symbol == "deflator") %>% 
  tq_gr(n = 1) %>% 
  filter(!is.na(Algeria))

tq_acf <- cal_factory(
  function(num) {
    acf(num, lag.max = 24, plot = FALSE)$acf
  },
  function(idx) {
    paste0("acf", 0:24)
  },
  function(itv) {
    "?"
  }
)

defl_gr1 %>% 
  tq_acf() %>% 
  filter(date == "acf12") %>% 
  pivot_longer(!c(date, symbol), names_to = "area") %>% 
  arrange(desc(value))


```

## cal_factory_zoo: function factory for calculation utilizing zoo package

You can utilize functions in zoo package in `cal_factory_zoo`. Here, for example, I utilize `zoo::rollmean`, and manufacture `tq_rollmean`, which has the same functionality of `tq_ma`. Although the output is not exactly the same due to floating point calculations, you can consider it as the same with `near` tolerance.

```{r cal_factory_zoo}
tq_rollmean <- cal_factory_zoo(
  function(num, ...) {
    zoo::rollmean(num, ...)
  },
  function(idx) idx,
  function(itv) itv
)

all_equal(
   tq_rollmean(eer_ts, k = 3, align = "right", fill = NA)[, "date"],
   tq_ma(eer_ts, n = 3)[, "date"]
)

near(tq_rollmean(eer_ts, k = 3, align = "right", fill = NA)$Algeria,
     tq_ma(eer_ts, n = 3)$Algeria) %>%
  all(na.rm = TRUE)


```

Speed is comparable. If you need speed, don't mind which function to choose, instead consider to spread to wide format.

```{r speed}
system.time(tq_ma(eer_ts, n = 3))
system.time(tq_rollmean(eer_ts, k = 3, align = "right", fill = NA))

eer_ts_long <- eer_ts %>% 
  pivot_longer(!c(date, symbol), names_to = "area")

system.time(tq_ma(eer_ts_long, n = 3))
system.time(tq_rollmean(eer_ts_long, k = 3, align = "right", fill = NA))


```

## cal_factory_ts: function factory for calculation utilizing ts class

`tq_sa` is manufactured by `cal_factory_ts` function. This function factory is only for interval "1Q" (quarterly) or "1M" (monthly). `tq_sa` is the only example for now.

```{r cal_factory_ts}
tq_sa <- cal_factory_ts(
  function(num_ts, ...) {
    num_ts %>%
      seasonal::seas(...) %>%
      seasonal::final() %>%
      as.numeric()
  },
  function(idx) {
    idx
  },
  function(itv) {
    itv
  }
)


```

## cal_factory_xts: function factory for calculation utilizing xts package

`xts::to.period` function can convert to OHLC (Open, High, Low, Close) in lower frequency. If OHLC = TRUE (default), one column expands to four, so you must input long format with only one numeric column. If you set OHLC = FALSE, you can input wide format, and get C (Close).

```{r cal_factory_xts}
tq_to_quarters <- cal_factory_xts(
  function(num, ...) {
    xts::to.period(num, period = "quarters", indexAt = "last", ...)
  },
  function(idx) {
    idx
  },
  function(itv) {
    "1Q"
  }
)

eer_ts %>% 
  select(date, symbol, Algeria) %>% 
  tq_to_quarters()

eer_ts %>% 
  tq_to_quarters(OHLC = FALSE)


```

EOL
